### 2021-03-18 ch/4-structure.tex
New motivating example for NIFR

As an example, consider a dataset where the task is to distinguish smiling from non-smiling faces.
Say, we initially have a very diverse and balanced training set,
but then, from the set of smiling faces, we remove nearly all samples where the person does not have red hair,
and from the set of non-smiling faces, we remove nearly all samples with non-black hair.
The result is that even though the labels are still correct, they do not point reliably to the right target anymore.
As hair colour is easier to recognise with a \ac{CNN}, the labels have effectively been turned into hair colour labels.
In other words, hair colour has become a \emph{spurious attribute}.
In the following, we denote the spurious attribute with \(s\),
as it takes a role that is very similar to that of the sensitive attribute that was also denoted by \(s\).
However, there is a difference in emphasis between a \emph{sensitive} and a \emph{spurious} attribute:
the former indicates that the attribute should not be used for legal or ethical reasons,
whereas the latter can be any attribute that is associated with the class label \(y\) in an undesired way
that leads to lower quality predictions.

>>>>

As an example, consider the scenario where someone wants to create a classifier
to distuinguish between sheep and cows, that is supposed to work anywhere on earth.
However, they take a shortcut while creating the dataset
and take all their sheep images from hot and dry countries
and all their cow images from mild and rainy countries.
In this case, the dataset is lacking cow images in dry landscapes,
and is lacking sheep from green landscapes;
the dataset exhibits a strong sampling bias.
The result is that even though the labels correctly correspond to cows and sheep,
they do not point reliably to the right target anymore.
As background colour is easier to recognise with a \ac{CNN} than animal species,
the labels have effectively been turned into landscape labels.
In other words, landscape has become a \emph{spurious attribute}.
%
% As an example, consider a dataset where the task is to distinguish smiling from non-smiling faces.
% Say, we initially have a very diverse and balanced training set,
% but then, from the set of smiling faces, we remove nearly all samples where the person does not have red hair,
% and from the set of non-smiling faces, we remove nearly all samples with non-black hair.
% The result is that even though the labels are still correct, they do not point reliably to the right target anymore.
% As hair colour is easier to recognise with a \ac{CNN}, the labels have effectively been turned into hair colour labels.
% In other words, hair colour has become a \emph{spurious attribute}.
In the following, we denote the spurious attribute with \(s\),
as it takes a role that is very similar to that of the sensitive attribute that was also denoted by \(s\).
However, there is a difference in emphasis between a \emph{sensitive} and a \emph{spurious} attribute:
the former indicates that the attribute should not be used for legal or ethical reasons,
whereas the latter can be any attribute that is associated with the class label \(y\) in an undesired way
that leads to lower quality predictions.
